```{r, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, message=FALSE, warning=FALSE, fig.align="center", fig.retina=3
)

old.hooks <- fansi::set_knit_hooks(knitr::knit_hooks)
options(crayon.enabled=TRUE)
```

# Lab 3 &mdash; September 29

## Some modern data science packages

The [Tidyverse](https://www.tidyverse.org/packages/) is a collection of packages for modern data
science. [Tidymodels](https://www.tidymodels.org/packages/) is another collection of packages that
extends the [tidyverse](https://www.tidyverse.org/packages/) to modelling. For these labs, I won't
be using the entire tidyverse and tidymodels. For now, I recommend only installing a few of their
component packages.

They are [tibble](https://tibble.tidyverse.org/), [dplyr](https://dplyr.tidyverse.org/),
[ggplot2](https://ggplot2.tidyverse.org/), and [broom](https://broom.tidymodels.org/). These
packages can be installed using:

```{r, eval=FALSE}
install.packages("tibble")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("broom")
```

After successfully installing these packages, you can load them using:

```{r}
library(tibble)
library(dplyr)
library(ggplot2)
library(broom)
```

Additionally, [ggplot2](https://ggplot2.tidyverse.org/) comes with various plotting themes (examples
of which can be found [here](https://ggplot2.tidyverse.org/reference/ggtheme.html#examples)). By
default, the plotting background is grey but I prefer a white background so I will also call:

```{r}
theme_set(theme_bw())
```

The [tibble](https://tibble.tidyverse.org/) package provides us with the usage of "tibbles" which
are an extension of the base-R data frame. Of its many features, we will mostly be using tibbles
for its [pretty-printing](https://tibble.tidyverse.org/articles/tibble.html#printing) features and
the fact that it integrates well with the other packages.

The [dplyr](https://dplyr.tidyverse.org/) package provides us with tools for data wrangling where
base-R equivalents can feel a bit clunky. Examples of data wrangling include: adding/removing
observations based on the presence/absence of one or more conditions, creating additional variables
in a data set, and condensing existing data into summaries.

The [ggplot2](https://ggplot2.tidyverse.org/) package is an alternative to base-R graphics
and has functions that are "smarter" than some that exist in base-R. For example, in the previous
lab, I mentioned that when plotting lines in base-R, points needed to sorted from left to right
since the `lines()` function joins points in the order that they appear. Without sorting, the
resulting line would appear jagged or unevenly coloured in certain areas. With
[ggplot2](https://ggplot2.tidyverse.org/), we have separate functions for joining points as they
appear in the data and joining points in order from left to right.

The [broom](https://broom.tidymodels.org/) package contains tools for cleaning up model output and
extracting model data with diagnostic information that is ready for visualisation with
[ggplot2](https://ggplot2.tidyverse.org/).


## Back to simple linear regression

### The `mpg` data set

Consider the fuel economy data set found in the [ggplot2](https://ggplot2.tidyverse.org/) package.
We can load this data set by calling:

```{r}
data(mpg, package="ggplot2")
```

Suppose we are interested in fitting a simple linear regression model with highway miles per gallon
(`hwy`) as the response variable and city miles per gallon (`cty`) as the predictor variable.
Before fitting the model, we should check to see if there is a linear relationship between our
chosen variables.

### Check for a linear relationship

To create this plot, we use the function `ggplot`.

```{r}
ggplot(mpg, aes(x=cty, y=hwy))+
  geom_point()
```

- The call to `ggplot` can be thought of as the initialisation of the drawing canvas
- The first argument of `ggplot` is the data set, `mpg`
- The second argument of `ggplot` is the mapping, i.e. *"which variables contain the data that I am
interested in plotting?"*
- The mapping is created by supplying the names of the variables found within your data set to the
aesthetics function, `aes()`.
- Finally, we can add (with a "plus" symbol) a layer of points with `geom_point()` that will inherit
any aesthetics supplied in the initialisation of the canvas
- In other words, while not specified within `geom_point()`, the points are drawn at positions
`x=cty` and `y=hwy`

From the plot above, there appears to be a linear relationship between the two variables. We can
begin fitting our linear model.


### Fitting the linear model

```{r}
lm_miles <- lm(hwy ~ cty, data=mpg)

summary(lm_miles)
```

From the regression output, we have that $\widehat{\beta}_{0} = 0.892$ and 
$\widehat{\beta}_{1} = 1.337$. The equation of the fitted line is:

\[\widehat{y}_{i} \,=\, 0.892 \,+\, 1.337x_{i}\]

We can interpret $\widehat{\beta}_{0} = 0.892$ as the mean highway miles per gallon when the city
miles per gallon has a value of zero. However, this doesn't really make sense in the context of our
data since you can't have a car that travels zero city miles on a gallon on gas.

We can interpret $\widehat{\beta}_{1} = 1.337$ as the mean change in highway miles per gallon per 
unit increase in city miles per gallon. As such, for a 1 mpg increase in city mpg, we expect a 1.337
mpg increase in highway mpg.


### Visualising the fit

Here is where [ggplot2](https://ggplot2.tidyverse.org/) and [broom](https://broom.tidymodels.org/)
shine! We first use `broom::augment()` on our linear model to extract the data that was used to fit
the model, along with additional diagnostics (such as the residuals).

```{r}
lm_miles_aug <- augment(lm_miles)

lm_miles_aug
```

In the first two columns of `lm_miles_aug`, we have recovered the data that we initially passed
into `lm()`. The `.fitted` column contains the fitted values of our model and the `.resid` column
contains the raw residuals.

The plot that we wish to build will use the newly created `lm_miles_aug` data set and will
have the following features:

- The base plot is a scatterplot with points located at `x=cty` and `y=hwy`
- The fitted line will be drawn on top with points located at `x=cty` and `y=.fitted`

Since the points and lines that we wish to draw no longer share a common y-aesthetic, we may not
want to declare a y-aesthetic in the initialisation of the canvas, and instead, declare the
y-aesthetic in the individual layers.

```{r}
ggplot(lm_miles_aug, aes(x=cty))+
  geom_point(aes(y=hwy))+
  geom_line(aes(y=.fitted), colour="#3366FF", lwd=1.5, alpha=0.6)
```

Notice that in building this plot, I did not sort any of the points! By default, `geom_line()`
connects points from left to right. I have added some additional arguments to `geom_line()` outside 
of the aesthetics since these values do not depend on the data:

- `colour` controls the colour of the line (can also use `color`)
- `lwd` controls the width of the line (can also use `linewidth`)
- `alpha` controls the transparency of the line

To see the construction of the above plot as "an addition of layers" we can run

```{r, eval=FALSE}
ggplot(lm_miles_aug, aes(x=cty))
```

followed by

```{r, eval=FALSE}
ggplot(lm_miles_aug, aes(x=cty))+
  geom_point(aes(y=hwy))
```

followed by

```{r, eval=FALSE}
ggplot(lm_miles_aug, aes(x=cty))+
  geom_point(aes(y=hwy))+
  geom_line(aes(y=.fitted), colour="#3366FF", lwd=1.5, alpha=0.6)
```


### Confidence and prediction intervals

#### Confidence intervals

Suppose we are interested in finding a point estimate and a 95% confidence interval for the mean
highway miles per gallon for vehicles with a city miles per gallon value of 20. We can obtain these
values by using the `predict()` function. More specifically, when we pass our `lm` object to
`predict()`, it actually calls the more specific `predict.lm()` function. To see what we need to get
our point estimate and confidence interval, let's pull up the associated documentation.

```{r, eval=FALSE}
?predict.lm
```

We will need to supply to `predict()`:

- A linear model object
- A data frame containing variables common to the linear model, and values upon which to predict
  - Note that if our model has predictor variable `cty`, we cannot supply
  `newdata = data.frame(x=20)`, we must supply `newdata = data.frame(cty=20)`
- If we want a confidence interval, we should specify `interval = "confidence"`
- The default confidence level is 0.95

```{r}
predict(
  lm_miles,
  newdata = data.frame(cty=20),
  interval = "confidence"
)
```

From the above output, the point estimate is 27.641. The lower bound of the 95% confidence interval
is 27.360 and the upper bound is 27.922. This says that with 95% confidence, for vehicles with a
city miles per gallon value of 20, the mean highway miles per gallon will range between 27.360 and
27.922.


#### Prediction intervals

Suppose we are now interested in finding a point estimate and a 95% prediction interval for the
highway miles per gallon of a vehicle that has a city miles per gallon of 20. We can obtain these
values in a manner similar to the previous example:

```{r}
predict(
  lm_miles,
  newdata = data.frame(cty=20),
  interval = "prediction"
)
```

From the above output, the point estimate is 27.641. The lower bound of the 95% prediction interval
is 24.177 and the upper bound is 31.105. This says that with 95% confidence, a vehicle with a city
miles per gallon value of 20 will have a highway miles per gallon value between 24.177 and 31.105.


#### Spot the differences

The point estimates in both cases were identical. However, the lower and upper bounds of the
confidence and prediction intervals were different. This is unsurprising since the formulas used to
compute the bounds of the two intervals are different.

It should be noted that, in general, for a fixed confidence level, the prediction interval will be
wider than its corresponding confidence interval.


### Correlations

Using the previously obtained augmented model data (`lm_miles_aug`), we can easily obtain the
correlations of

- $X_{i}$ and $Y_{i}$,
- $Y_{i}$ and $\widehat{Y}_{i}$,
- $X_{i}$ and $\widehat{Y}_{i}$,

compare them among each other, and compare them to the model's coefficient of determination,
$R^{2}$. Correlations are obtained using the `cor()` function. We will also use the `summarise()`
function from the [dplyr](https://dplyr.tidyverse.org/) package to help us obtain the correlations
while staying in context of our augmented data so that we can reference our variable using bare
names (i.e. we don't need to use `$` to reference our variables).

```{r}
summarise(
  lm_miles_aug,
  corr_x_y = cor(cty, hwy),
  corr_y_fitted = cor(hwy, .fitted),
  corr_x_fitted = cor(cty, .fitted)
)
```

Note also that:

```{r}
summarise(
  lm_miles_aug,
  corr_x_y_squared = cor(cty, hwy)^2,
  corr_y_fitted_squared = cor(hwy, .fitted)^2,
  corr_x_fitted_squared = cor(cty, .fitted)^2,
  r.squared = summary(lm_miles)$r.squared
)
```

Are these values mathematically related or is this coincidence? &#x1F632; *(See assignment 2
question 3).*


## Simple linear regression with transformations

Talk about this next time
