```{r, include=FALSE}
knitr::opts_chunk$set(
  echo=TRUE, message=FALSE, warning=FALSE, fig.align="center", dev="svglite"
)

old.hooks <- fansi::set_knit_hooks(knitr::knit_hooks)
options(crayon.enabled=TRUE)
```

# Lab 9 &mdash; December 1

## Packages

```{r}
library(tibble)
library(dplyr)
library(ggplot2)
theme_set(theme_bw())
```


## Logistic regression with binomial ($n > 1$) data

**Note**: For your assignment, you should use the code provided in the `logitfit1` file on
Brightspace. However, the code and output shown below are equivalent and also fitted via
iteratively reweighted least squares (IRWLS).


### Shocking some cows

Consider the data below, that is similar to Table 12.3 in the handout, but with the number of
responses in the first experiment changed from 0 to 2. Note that the code in the `logitfit1` file
uses a value of 0.0001 rather than 0 to prevent the resulting model deviance from being `NaN`.

| Current (mA) | Trials | Responses | Proportion of Responses |
|:------------:|:------:|:---------:|:-----------------------:|
| 0            | 70     | 2         | 0.029                   |
| 1            | 70     | 9         | 0.129                   |
| 2            | 70     | 21        | 0.300                   |
| 3            | 70     | 47        | 0.671                   |
| 4            | 70     | 60        | 0.857                   |
| 5            | 70     | 63        | 0.900                   |

For the $i^{\text{th}}$ experiment:

- $x_{i}$ is the amount of current applied
- $n_{i}$ is the number of trials
- $y_{i}$ is the number of responses
- $n_{i} - y_{i}$ is the number of no-responses
- $y_{i} / n_{i}$ is the proportion of responses

We wish to fit the model:

\[E\left(\frac{y_{i}}{n_{i}}\right) \,=\, \theta_{i} \,=\, \frac{\exp{(\beta_{0} + \beta_{1}x)}}
{1 + \exp{(\beta_{0} + \beta_{1}x)}}\]

Inputting the data into R:

```{r}
cows <- tibble(
  current = 0:5,
  trials = rep(70, 6),
  responses = c(2, 9, 21, 47, 60, 63),
  proportion = responses / trials
)

cows
```


### Fitting the model

When fitting the model, we represent the dependent variable as a matrix of successes and failures.
As such the representation of our dependent variable will look like:

```{r}
with(cows, cbind(responses, no_responses = trials - responses))
```

Fitting the model now:

```{r}
full_model <- glm(cbind(responses, trials - responses) ~ current, family=binomial, data=cows)
```

Note that in the `family` argument, we specify `binomial` but we do not need to wrap it in quotes
as it is a special object.

```{r}
summary(full_model)
```

This gives us the fitted equation on the predictor scale (log-odds):

\[\widehat{\eta}_{i} \,=\, \widehat{\beta}_{0} + \widehat{\beta}_{1}X_{i} \,=\, -3.1020 + 1.1837X_{i}\]

To transform it onto the response scale (probabilities), we use the relationship that:

\[\widehat{\eta}_{i} \,=\, g(\widehat{\theta}_{i}) \,=\, \log{\left(\frac{\widehat{\theta}_{i}}{1 - \widehat{\theta}_{i}}\right)}\]

thus,

\[\widehat{\theta}_{i} \,=\, g^{-1}(\widehat{\eta}_{i}) \,=\, \frac{\exp{(\widehat{\eta}_{i})}}{1 + \exp{(\widehat{\eta}_{i})}}\]


### Obtaining the predicted probabilities

As with the OLS linear models, we can pass a GLM into the `predict()` function which will then
call `predict.glm()`.

```{r}
predict(full_model)
```

However, notice that that some of these values are outside of $[0,1]$! This is because these are
the predicted values on the predictor scale (log-odds). If we want the predicted values on the
response scale (probabilities), from the documentation (`?predict.glm`), we need to specify that
`type="response"`.

```{r}
predict(full_model, type="response")
```


### Visualising the fit

Since our data is quite small, unlike with straight line fits where we can just proceed to connect
the dots and still obtain a straight line, here, we need to supplement our data with a sufficient
number of grid points to get as close as possible to the true shape of the predicted curve.

```{r}
fitted_vals <- full_model %>%
  predict(newdata = tibble(current=seq(0, 5, 0.01)), type="response") %>%
  as_tibble_col(column_name=".fitted") %>%
  add_column(current=seq(0, 5, 0.01), .before=".fitted")

fitted_vals
```

```{r}
ggplot(data=NULL, aes(x=current))+
  geom_point(data=cows, aes(y=proportion))+
  geom_line(data=fitted_vals, aes(y=.fitted), colour="#3366FF", lwd=1.5, alpha=0.6)+
  labs(x="Current (mA)", y="Proportion of responses to electric shock")
```


### Analysis of deviance

to do: deviance, log-likelihoods, degrees of freedom, test statistic

```{r, eval=FALSE}
deviance(full_model)

anova(full_model, test="Chisq")

reduced_model <- update(full_model, . ~ 1)

anova(reduced_model, full_model, test="Chisq")
```



